---
title: 饼图中的南辕北辙
author: Zhuoer Dong
date: '2018-05-20'
slug: 饼图-南辕北辙
tags: []
---

```{r knitr-setup, include=FALSE}
knitr::opts_chunk$set(message = FALSE, collapse = TRUE)
```

```{r install-pkg, message=FALSE, eval=Sys.getenv('CI') == 'true'}
# install needed R packages
remotes::update_packages(c('magrittr', 'tidyverse', 'cowplot'), upgrade = TRUE)
```


# Beginning

为了展示毕业论文的数据，我想要画个饼图。`pie()` 虽然简单，但我肯定要用高大上的“哥哥作图”（**ggplot2**）。

```{r}
library(magrittr)
library(ggplot2)

df <- tibble::tribble(
    ~type, ~n,
    'GDS', 1772,
    'other GSE', 18309
) %>% dplyr::mutate(percent = n/sum(n), label = scales::percent(percent))

df
```


# Development

Althought I already know pie chart is a variant of bar plot with polar coordinate, I don't know the precise code. As usual, Stack Overflow provides a good [example](https://stackoverflow.com/questions/45657990#answer-45659764).

I applied the code, but the order of label seems to be wrong:


```{r}
ggplot(df) +
  geom_bar(aes(x = "", y = percent, fill = type), stat="identity", width = 1) +
  coord_polar("y", start = 0) +
  theme_void() +
  geom_text(aes(x = 1, y = cumsum(percent) - percent/2, label = label))
```

After a while, I find a solution:

```{r}
ggplot(df) +
  geom_bar(aes(x = "", y = percent, fill = factor(type, levels = rev(type))), stat="identity", width = 1) +
  coord_polar("y", start = 0)+
  theme_void()+
  geom_text(aes(x = 1, y = cumsum(percent) - percent/2, label = label))
```


# Climax

However, I still don't understand what happens, so I start to explore it step by step.

I refered to the [official documentation](https://ggplot2.tidyverse.org/reference/coord_polar.html) for how pie chart is evolved from bar plot. 

To understand why the order of label is incorrect, we need to examine what happens in the original bar plot:

```{r fig.dim=c(12, 3), out.extra='style="width:100%;"'}
cowplot::plot_grid(
    ggplot(df) + 
        geom_col(aes(type, percent, fill = type)) + 
        geom_text(aes(x=type, y =  percent/2, label=type)),
    ggplot(df) + 
        geom_col(aes('', percent, fill = type)) + 
        geom_text(aes(x='', y = cumsum(percent) - percent/2, label=type)),
    ggplot(dplyr::arrange(df, dplyr::desc(type)))+ 
        geom_col(aes('', percent, fill = type)) + 
        geom_text(aes(x='', y = cumsum(percent) - percent/2, label=type)),
    nrow = 1
)        
```

- y of label is determined by the row order in data, i,e, the earlier a row appears, the smaller y value, and the lower position. 
- The order of bar is determined by factor level order (default is lexicographic order of level strings). The smaller level order the higher position.

So `dplyr::arrange(..., dplyr::desc(...))` can make top row have small y value and big level order, thus sovle the problem ^[`factor(type, levels = rev(type))` is a coincidence, it only works in this case.].


# Afterword

When I come back to polish this post on 2019/03/03, I  ponder the official documentation for half an hour and realize I still haven't fully understand the transition from bar plot to pie chart. 

I think the documentation is not quiet clear. The most confusing thing is how radius is determined, since it's neither mentioned nor deducible from examples.

The examples have two drawbacks:
 
1. the yuse real data, so reader can't know the exact value of y.
2. the corresponding bar plot is not shown, so reader have to imagine it.

Therefore, I write the following examples, hoping they can save time when I forget how pie chart is drawn the next time:

```{r fig.dim=c(9, 3)}
tbl <- tibble::tribble(
    ~class, ~value,
    'A', 1,
    'B', 2,
    'C', 4
)

bar <- ggplot(tbl) + geom_col(aes(class, value), width = 1)
cowplot::plot_grid(
    bar,
    bar + coord_polar(theta = 'x'),
    bar + coord_polar(theta = 'y'),
    nrow = 1
)
```

From the aboving plots, we can know that:

- `theta = 'x'`: x becomes angle, y becomes radius
- `theta = 'y'`: x becomes radius, y becomes angle

```{r fig.dim=c(9, 3)}
stack <- ggplot(tbl) + geom_col(aes('1', value, fill = class), width = 1)
cowplot::plot_grid(
    stack,
    stack + coord_polar(theta = 'x'),
    stack + coord_polar(theta = 'y'),
    nrow = 1
)
```

In the aboving plots, `fill` divide y value:

- When y is radius, we have concentric circles.
- When y is angle, we have sectors.

