---
title: Yet another dark magic in R, empty argument
author: Zhuoer Dong
date: '2018-05-20'
slug: r-empty-argument
categories: 2018
tags: []
authors: []
---

```{r knitr-setup}
knitr::opts_chunk$set(error = TRUE, collapse = TRUE)
```

```{r install-pkg, message=FALSE, eval=Sys.getenv('CI') == 'true'}
# install needed R packages
remotes::update_packages(c('magrittr', 'testthat', 'tibble'), upgrade = TRUE)
remotes::install_github(c('krlmlr/deparse'))
```


# Beginning

在用 testthat 的时候发现了 `dput()` 真的很好用 [^dput-testthat]，就想找一下有没有把 tibble 转化成 `tribble(...)` 的函数。

[^dput-testthat]: 例如你要检查 `x` 是不是期望的值，`dput()` 就可以告诉你什么样的 R 代码可以构造出期望的值：

    ```{r}
    x <- list()
    x$a = 1:3
    x$b = c('a', 'b', 'c')
    
    dput(x)

    testthat::expect_identical(x, list(a = 1:3, b = c("a", "b", "c")))
    testthat::expect_identical(x[1], list(a = 1:3, b = c("a", "b", "c")))
    ```

结果 krlmlr 大人已经 [实现](https://github.com/tidyverse/tibble/issues/127) 了:

```{r}
(df <- tibble::tibble(x = 1:2, y = 2:1))

deparse::deparsec(df, as_tribble = TRUE)
```


# Development

继续逛了一下，看到有 [好事者](https://github.com/krlmlr/deparse/issues/6) 还是不满意，觉得

```r
tribble(
  ~mpg, ~cyl, ~disp, ~hp, ~drat, ~wt,   ~qsec, ~vs, ~am, ~gear, ~carb,
  21,   6,    160,   110, 3.9,   2.62,  16.46, 0,   1,   4,     4,    
  21,   6,    160,   110, 3.9,   2.875, 17.02, 0,   1,   4,     4    
)
```

应该变成

```r
tibble::tribble(
    ~mpg, ~cyl, ~disp, ~hp, ~drat,   ~wt, ~qsec, ~vs, ~am, ~gear, ~carb
  ,   21,    6,   160, 110,   3.9,  2.62, 16.46,   0,   1,     4,     4
  ,   21,    6,   160, 110,   3.9, 2.875, 17.02,   0,   1,     4,     4
)
```

他说的倒很有道理，后者使得调换不同的行变得很方便（前者把第1、2 行互换的话，需要给新的第1行加个`,`，再从新的第2行删个`,`）。但是 krlmlr 大人却[提出](https://www.bountysource.com/issues/48133299-suggested-function-fake_tribble):

```r
tribble(
  ~a, ~b,
   1,  2,
)
```

# Climax

这就很有意思了，不会造成语法错误吗？

Thus I have a try myself, and find that empty argument means using parameter's default value:

```{r}
foobar <- function(a=1,b=2) {
	print(a)
	print(b)
}

foobar(3,)
foobar(,4)
foobar(3,4,)
```

Then I recall that I shouldn't feel strange for this syntax:

```r
mtcars[1, ]
mtcars[ , ]
```


# Epilogue

Understanding what happens, I start to think maybe I can write `read_csv('', , 'cc')` instead of `read_csv('', T, 'cc')`.

No, absolutely not!

If I really don't want to type the `T`, I'd better write `read_csv('', col_types = 'cc')`.

> Type 10 more character now, save a lot of times in the future.



















