---
title: use datamodelr to create database schema diagrams
author: Zhuoer Dong
date: '2018-11-10'
slug: use-datamodelr-show-database-schema
tags: []
---

Thanks to _R for Data Science Solutions_ for telling me how to draw database schema diagrams.

Here we use the **datamodelr** package: [^dependencies-T]

[^dependencies-T]:
    As of this writing, the package use `Suggests` to list dependencies, I have lanched an [issue](https://github.com/bergant/datamodelr/issues/8). 
    
    But for now, we can only use `dependencies = T` to install "Depends", "Imports", "LinkingTo" and "Suggests".

```{r install-pkg, message=FALSE, eval=Sys.getenv('CI') == 'true'}
# install needed R packages
remotes::update_packages(c('magrittr', 'Lahman', 'DiagrammeR'), upgrade = TRUE)
remotes::install_github(c("bergant/datamodelr"))
```

Following is a simple example:

```{r fig.height=8, fig.width=6}
library(magrittr)
library(datamodelr)

list(Batting = Lahman::Batting, Master = Lahman::Master, Salaries = Lahman::Salaries) %>% 
    dm_from_data_frames() %>%
    dm_set_key("Batting", c("playerID", "yearID", "stint")) %>%
    dm_set_key("Master", "playerID") %>%
    dm_set_key("Salaries", c("yearID", "teamID", "playerID")) %>%
    dm_add_references(
        Batting$playerID == Master$playerID, 
        Salaries$playerID == Master$playerID
    ) %>% 
    dm_create_graph(rankdir = "LR", columnArrows = TRUE) %>% 
    dm_render_graph()
```

You can see more examples in the package's [GitHub homepage](https://github.com/bergant/datamodelr).

