---
title: '`ggplot2::ggplot_build()` abandans the dark magic I worked pass midnight'
author: Zhuoer Dong
date: '2018-12-01'
slug: ggplot-build-nullity-a-dark-magic
categories: 2018
tags: []
authors: []
---


```{r install-pkg, message=FALSE, eval=Sys.getenv('CI') == 'true'}
# install needed R packages
remotes::update_packages(c('gginnards', 'tibble', 'ggplot2'), upgrade = TRUE)
```


# Beginning

When I do rotation in Lu Lab, I'm assigned to write some R code for matrix process and plot part of exSeek project. In the process, I want to reproduce a plot in R, i.e.,  the second plot [here](https://github.com/dongzhuoer/lulab-rotation-summary/blob/master/exseek/matrix-process.md#show-variance-of-reference-genes).



# Development

The problem is that I want to label the text above each violin, but there isn't a easy way to get the highest y coordinate. `median()` is okay, but calculate the value by something like `median() + 1.5 * quantile(...)` is too complicated and ugly. The value has already been calculated in the plot, why we bother to do it again. ^[In fact, when my collaborator creates the plot in Python, he use `median()` plus a constant.]

After inspect the ggplot object (treat it as a list) for a while, I find a package, **gginnards**. It can show you the internal aesthetics computed by ggplot, like [this one](https://docs.r4photobiology.info/gginnards/articles/user-guide-1.html#data-input-to-stats) 



# Climax

However, it can only _print_ the value, while I want to _get_ it. So I write the following function (I worked pass 0:00, maybe close to 01:00):

```{r}
#' @title Get computed variables of a ggplot object
#'
#' @description
#' Access the computed variables, like `density` of [ggplot2::geom_density()]
#'
#' @param ggplot_obj ggplot object. [ggplot2::ggplot()] must contains
#'
#' @return tibble.
#'
#' @keywords internal
#'
#' @examples
#' get_ggplot_data(ggplot2::ggplot(mtcars, ggplot2::aes(wt, mpg)) + ggplot2::geom_point())
#' \donotrun{
#' # ggplot() must contain default x, y aesthesic.
#' get_ggplot_data(ggplot2::ggplot(mtcars) + ggplot2::geom_point(ggplot2::aes(wt, mpg)))
#' # failed with geom_density
#' }

get_ggplot_data <- function(ggplot_obj) {
	temp_file <- tempfile();
	on.exit(file.remove(temp_file))

	env <- new.env();
	assgin_to_data <- function(x, env) {print(x);env$data <- tibble::as_tibble(x)}

	ggplot_obj2 = ggplot_obj +
		gginnards::stat_debug_panel(
			summary.fun = assgin_to_data,
			summary.fun.args = list(env = env)
		)

	sink(temp_file)
	png(temp_file)
	print(ggplot_obj2)
	dev.off()
	sink()

	env$data
}
```

Let's have a try

```{r}
ggplot_obj <- ggplot2::ggplot(mtcars, ggplot2::aes(wt, mpg)) + ggplot2::geom_point()
```

```{r echo = F}
plot(ggplot_obj)
```

```{r}
get_ggplot_data(ggplot_obj)
```


# Epilogue

When I tidy the code in the morning today, I find that **ggplot2** actually provides an official interface for this purpose, `ggplot2::ggplot_build()`. Bang, my heart crashes.

As usual, I move the code here for memorial.
